---
title: "Modern Approaches to Difference-in-Differences"
subtitle: "Session 2: Including Covariates in the Parallel Trends Assumption"
format: clean-revealjs
author:
  - name: Brantly Callaway
    email: brantly.callaway@uga.edu
    affiliations: University of Georgia
knitr:
  opts_chunk:
    echo: true
bibliography:
  - refs.bib
  - labor_refs.bib
---

```{r echo=FALSE}
library(revealEquations)
```

## Outline

1. Introduction to Difference-in-Differences

2. [Including Covariates in the Parallel Trends Assumption]{.alert}

3. Common Extensions for Empirical Work

4. Dealing with More Complicated Treatment Regimes

5. Alternative Identification Strategies

## Covariates in the Parallel Trends Assumption

Including covariates in the parallel trends assumption can often make DID identification strategies more plausible:

. . .

[Example]{.alert-blue}

* [Minimum wage example:]{.alert} path of teen employment may depend on a state's population / population growth / region of the country

* [Job displacement example:]{.alert} path of earnings may depend on year's of education / race / occupation

. . .

However, there are a number of new issues that can arise in this setting...

$\newcommand{\E}{\mathbb{E}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\var}{\mathrm{var}}
\newcommand{\cov}{\mathrm{cov}}
\newcommand{\Var}{\mathrm{var}}
\newcommand{\Cov}{\mathrm{cov}}
\newcommand{\Corr}{\mathrm{corr}}
\newcommand{\corr}{\mathrm{corr}}
\newcommand{\L}{\mathrm{L}}
\renewcommand{\P}{\mathrm{P}}
\newcommand{\independent}{{\perp\!\!\!\perp}}
\newcommand{\indicator}[1]{ \mathbf{1}\{#1\} }$


## Plan: Covariates in the Parallel Trends Assumption

<br>

1. Identification with Two Periods

2. Limitations of TWFE Regression

3. Alternative Estimation Strategies

4. Multiple Periods

5. Minimum Wage Application

6. Dealing with "Bad" Controls

## Notation

Start with the case with only two time periods

Only need a little bit of new notation here:

* $X_{i,t=2}$ and $X_{i,t=1}$ &mdash; time-varying covariates

* $Z_i$ &mdash; time-invariant covariates

## Covariates in the Parallel Trends Assumption

<br>

::: {.callout-note}

### Conditional Parallel Trends Assumption


$$\E[\Delta Y(0) | X_{t=2}, X_{t=1},Z,D=1] = \E[\Delta Y(0) | X_{t=2}, X_{t=1},Z,D=0]$$
:::

::: {.fragment}

In words: Parallel trends holds conditional on having the same covariates $(X_{t=2},X_{t=1},Z)$.

:::

::: {.fragment}

[Minimum wage example]{.alert} (e.g.) Parallel trends conditional on counties have the same population (like $X_{t}$) and being in the same region of the country (like $Z$)

:::

```{r echo=FALSE, results="asis"}

title <- "Identification under Conditional Parallel Trends"

before <- "Under conditional parallel trends, we have that"

eqlist <- list("ATT &= \\E[\\Delta Y | D=1] - \\E[\\Delta Y(0) | D=1] \\hspace{150pt}",
               "&=\\E\\Big[ \\E[\\Delta Y | X,D=1] \\Big| D=1\\Big] - \\E\\Big[ \\E[\\Delta Y(0) | X, D=1] \\Big| D=1\\Big]",
               "&=\\E\\Big[ \\underbrace{\\E[\\Delta Y | X,D=1] - \\E[\\Delta Y(0) | X, D=0]}_{=:ATT(X)} \\Big| D=1\\Big]",
               "&= \\E[\\Delta Y | D=1] - \\E\\Big[ \\underbrace{\\E[\\Delta Y(0) | X, D=0]}_{=:m_0(X)} \\Big| D=1\\Big]")


after <- "

. . .

Intuition: (i) Compare path of outcomes for treated group to (conditional on covariates) path of outcomes for untreated group, (ii) adjust for differences in the distribution of covariates between groups.

. . .

This argument also require an [overlap condition]{.alert}

* For all possible values of the covariates, $p(x) := \\P(D=1|X=x) < 1$.

* In words: for all treated units, we can find untreated units that have the same characteristics

"

step_by_step_eq(eqlist, before, after, title)
```

```{r echo=FALSE, results="asis"}

title <- "Alternative Identification Strategy: Covariate Balancing"

before <- "[Example:]{.alert} Momentarily, suppose that the distribution of $X$ was the same for both groups, then

. . .

"

eqlist <- list("ATT &= \\E[\\Delta Y | D = 1] - \\E[\\Delta Y(0) | D = 1] \\hspace{150pt}",
  "&= \\E[\\Delta Y | D = 1] - \\E\\Big[ \\E[\\Delta Y(0) | X, D=0 ] \\Big| D=1\\Big]",
  "&= \\E[\\Delta Y | D = 1] - \\E\\Big[ \\E[\\Delta Y(0) | X, D=0 ] \\Big| D=0\\Big]",
  "&= \\E[\\Delta Y | D = 1] - \\E[\\Delta Y(0) | D=0]")


after <- "

. . .

$\\implies$ (even under conditional parallel trends) we can recover $ATT$ by just directly comparing paths of outcomes for treated and untreated groups."

step_by_step_eq(eqlist, before, after, title)
```

```{r echo=FALSE, results="asis"}

title <- "Alternative Identification Strategy: Covariate Balancing"

before <- "[More generally:]{.alert} We would not expect the distribution of covariates to be the same across groups.

However the idea of covariate balancing is to come up with [balancing weights]{.alert} $\\nu_0(X)$ such that the distribution of $X$ is the same in the untreated group as it is in the treated group after applying the balancing weights.  Then we would have that

. . .

"

eqlist <- list("ATT &= \\E[\\Delta Y | D=1] - \\E[\\Delta Y(0) | D=1] \\hspace{150pt}",
  "&= \\E[\\Delta Y | D=1] - \\E\\Big[ \\E[\\Delta Y(0) | X, D=0 ] \\Big| D=1\\Big]",
  "&= \\E[\\Delta Y | D=1] - \\E\\Big[ \\nu_0(X) \\E[\\Delta Y(0) | X, D=0 ] \\Big| D=0\\Big]",
  "&= \\E[\\Delta Y | D=1] - \\E[\\nu_0(X) \\Delta Y(0) | D=0]")


after <- "

. . .

$\\implies$ We can recover $ATT$ by re-weighting the untreated group to have the same distribution of covariates as the treated group has...and then just average"

step_by_step_eq(eqlist, before, after, title)
```

## Discussion

The arguments about suggest that, in order to estimate the $ATT$, we will either need to

1. Correctly model $\E[\Delta Y(0) | X, D=0]$ (i.e, specifiy a model for $m_0(X)$)

2. Balance the distribution of $X$ to be the same for the untreated group relative to the treated group.

. . .

Next, we will discuss how well this works for TWFE regressions and then alternative (more direct) estimation strategies.

# Limitations of TWFE Regressions {visibility="uncounted"}

## Limitations of TWFE Regressions

In this setting, it is common to run the following TWFE regression:

$$Y_{it} = \theta_t + \eta_i + \alpha D_{it} + X_{it}'\beta + e_{it}$$

::: {.fragment}

However, there are a number of issues: <!--(most of these apply even in the friendly setting with two periods)-->

[Issue 1:]{.alert} Issues related to multiple periods and variation in treatment timing still arise

:::

::: {.fragment}


[Issue 2:]{.alert} Hard to allow parallel trends to depend on time-invariant covariates

:::

::: {.fragment}


[Issue 3:]{.alert} Hard to allow for covariates that could be affected by the treatment

:::

::: {.fragment}

[Issues 4 & 5:]{.alert} (harder to see) Can perform poorly for including time-varying covariates in the parallel trends assumption

:::

## Side-Discussion: Interpreting regressions under unconfoundedness

Consider the cross-sectional regression
$$
  Y_i = \alpha D_i + X_i'\beta + e_i
$$
where we assume uncoundedness: $Y(0) \independent D | X$

. . .

[View 1:]{.alert} Correctly specified model for $\E[Y|X,D]$.

* This implies a model for $m_0(X)$: $m_0(X) = X'\beta$

* It also restricts treatment effect heterogeneity:

  For any $X$, $\E[Y|X,D=1] - \E[Y|X,D=0] = \alpha$ $\implies$ treatment effects don't systematically vary with $X$.

* see, e.g., @chaisemartin-ciccia-dhaultfoeuille-knau-2024 for related linearity tests in the context of TWFE

## Side-Discussion: Interpreting regressions under unconfoundedness {visibility="uncounted"}

Consider the cross-sectional regression
$$
  Y_i = \alpha D_i + X_i'\beta + e_i
$$
where we assume uncoundedness: $Y(0) \independent D | X$

[View 2:]{.alert} Linear model as approximation to possibly more complicated conditional expectation (e.g., @angrist-1998)

. . .

You can show some interesting/useful results in this case &nbsp; $\rightarrow$

## Side-Discussion: Interpreting regressions under unconfoundedness

1. [$\alpha$ can be re-interpreted as a weighting estimator]{.alert}

    $$ \alpha = \E\Big[w_1(X) Y \Big| D=1\Big] - \E\Big[w_0(X) Y \Big| D=0\Big] $$

    where
    $$ w_1(X) := \frac{\big(1-\L(D|X)\big) \pi}{\E\big[(D-\L(D|X))^2\big]} \ \ \textrm{and} \ \ w_0(X) := \frac{\L(D|X)(1-\pi)}{\E\big[(D-\L(D|X))^2\big]} $$

    and the result here follows (basically) immediately using FWL/partialling out arguments.

## Side-Discussion: Interpreting regressions under unconfoundedness {#interpreting-regressions}

2. [$\alpha$ is equal to a weighted average of $ATT(X)$ plus misspecification bias terms:]{.alert}

    $$ \alpha = \E\Big[w_1(X) ATT(X) \Big| D=1\Big] - \E\Big[w_1(X)\Big(\E[Y|X,D=0] - \L_0(Y|X)\Big) \Big| D=1\Big] $$

    The misspecification bias component is equal to 0 if either:

    * $\E[Y|X,D=0] = \L_0(Y|X)$ (model for untreated potential outcomes is linear in $X$)

    * The implicit regression weights are covariate balancing weights; i.e. for any function of the covariates $g$

        $$ \E\Big[ w_1(X) g(X) \Big| D=1\Big] = \E\Big[ w_0(X) g(X) \Big| D=1 \Big]$$

        [[more details](#reg-weights-more-details)]


## Limitations of TWFE Regressions

Let us now return to the TWFE regression:
$$Y_{it} = \theta_t + \eta_i + \alpha D_{it} + X_{it}'\beta + e_{it}$$
and specialize to the case with two time periods, so that we ultimately run the regression

$$\Delta Y_{it} = \Delta \theta_t + \alpha D_{it} + \Delta X_{it}'\beta + \Delta e_{it}$$

. . .

Using the same arguments as above, you can show that:
$$ \alpha = \E\Big[ w_1(\Delta X) ATT(X_{t=2},X_{t=1},Z)\Big| D=1 \Big] + \E\Big[ w_1(\Delta X) \Big( \E[\Delta Y | X_{t=2}, X_{t=1}, Z] - \L_0(\Delta Y | \Delta X) \Big) \Big| D=1 \Big]$$

. . .

Similar to above, $\alpha$ is equal to weighted averages of:

* conditional-on-covariates $ATT$'s

* misspecification bias

## Limitations of TWFE Regressions

The misspecification bias term is equal to 0 if:

* $\E[\Delta Y | X_{t=2}, X_{t=1}, Z] = \L_0(\Delta Y | \Delta X)$.  This amounts to:
    * A condition about linearity (makes sense...)
    * Changing the identification strategy from one where parallel trends only depends on $\Delta X$ rather than on $X_{t=1}$, $X_{t=2}$ and $Z$.

* The implicit regression weights $w_1(\Delta X)$ and $w_0(\Delta X)$ balance the distribution of $(X_{t=2}, X_{t=1}, Z)$ for the treated group relative to the untreated group.


## Limitations of TWFE Regressions

In @caetano-callaway-2023, we refer to the misspecification bias term above as [hidden linearity bias]{.alert}.

. . .

What we mean is that the implications of a linear model may be much more severe in a panel data setting than in the cross-sectional setting:

* It effectively changes the identification to one where what matters is changes in covariates over time

* The regression will balance (in mean) terms that show up in the regression (i.e. $\Delta X$), but it won't balance terms that don't show up (e.g., $Z$) or other terms such as $X_{t=1}$ or $X_{t=2}$.

## Limitations of TWFE Regressions

[How much does this matter in practice?]{.alert}

* Not all that easy to check how far away $\E[\Delta Y | X_{t=1}, X_{t=2}, Z, D=0]$ is from $\L_0(\Delta Y|\Delta X)$

. . .

* Instead, an easier idea is to apply implicit regression weights to $Z$, $X_{t=1}$, etc.:

    $$ \E\left[w_1(\Delta X) \begin{pmatrix} Z \\ X_{t=1} \end{pmatrix} \middle| D=1 \right] \overset{?}{=} \E\left[w_0(\Delta X) \begin{pmatrix} Z \\ X_{t=1} \end{pmatrix} \middle| D=0 \right] $$

    which gives us a way to diagnose the sensitivity of the TWFE regression to hidden linearity bias.
    * This is easy to check in practice: weights just depend on linear projections that are easy to directly estimate
    * If these are close, it suggests that misspecification bias is small.
    * If not, then it matters a lot whether or not the model is correctly specified.

## Limitations of TWFE Regressions

[Issue 5:]{.alert} Even if none of the previous 4 issues apply, $\alpha$ is a weighted average of $ATT(X)$.  However,

. . .

* The weights can be negative

* The weights suffer from *weight reversal* (e.g., @sloczynski-2022):
  * Too much weight on $ATT(X)$ for values of the covariates that are relatively uncommon for the treated group relative
  * Too little weight on $ATT(X)$ for values of the covariates that are relatively common for the treated group


# Alternative Estimation Strategies {visibility="uncounted"}

## Regression Adjustment (RA)

<!--The challenging term to deal with in the previous expression for $ATT$ is-->

Recall our first identification result above:

$$ATT = \E[\Delta Y | D=1] - \E\Big[ \underbrace{\E[\Delta Y(0) | X, D=0]}_{=:m_0(X)} \Big| D=1\Big]$$

The most direct way to proceed is by proposing a model for $m_0(X)$.  For example, $m_0(X) = X'\beta_0$.

. . .

* Notice that linearity of untreated potential outcomes is exactly the same condition we needed for the TWFE regression to be a weighted average of conditional-on-covariates $ATT$'s.

. . .

* However, here $X$ includes $(X_{t=2}, X_{t=1}, Z)$ rather than only $\Delta X$.
  - This means that there could still be misspecification bias, but there is no hidden linearity bias

. . .

* Moreover, if linearity holds, we directly target $ATT$, rather than recovering a hard-to-interpret weighted average of $ATT(X)$.

. . .

* $\implies$ there is a strong case to go with RA as a default option over TWFE.

## Regression Adjustment (RA) {visibility="uncounted"}

Recall our first identification result above:

$$ATT = \E[\Delta Y | D=1] - \E\Big[ \underbrace{\E[\Delta Y(0) | X, D=0]}_{=:m_0(X)} \Big| D=1\Big]$$


This expression suggests a [regression adjustment]{.alert} estimator:

$$ATT = \E[\Delta Y | D=1] - \E[X'\beta_0|D=1]$$

. . .

and we can estimate the $ATT$ by

* Step 1: Estimate $\beta_0$ using untreated group

* Step 2: Compute predicted values for treated units: $X_i'\hat{\beta}_0$

* Step 3: Compute $\widehat{ATT} = \displaystyle \frac{1}{n} \sum_{i=1}^n \frac{D_i}{\hat{\pi}} \Delta Y_i - \frac{1}{n} \sum_{i=1}^n \frac{D_i}{\hat{\pi}} X_i'\hat{\beta}_0$

## Side-Discussion on Imputation Estimators

(One shot) imputation estimators that include covariates typically involve estimating the model
$$Y_{it}(0) = \theta_t + \eta_i + X_{it}'\beta + e_{it}$$

. . .

[Issue 1:]{.alert} Multiple periods <span style="color:green">&#10004;</span>

[Issue 2:]{.alert} Time-invariant covariates <span style="color:orange;font-weight:bold">&#9888;</span>

[Issue 3:]{.alert} Covariates affected by the treatment &#10060;

[Issue 4:]{.alert} Hidden linearity bias &#10060;

* You can see that it implicitly relies on $\E[\Delta Y(0) | X_{t=2}, X_{t=1}, Z, D=0] = \E[\Delta Y(0) | \Delta X, D=0]$

[Issue 5:]{.alert} Weighted average of $ATT(X)$ <span style="color:green">&#10004;</span>

## Propensity Score Weighting (IPW)

Alternatively, recall our identification strategy based on re-weighting:
$$ ATT = \E[\Delta Y | D=1] - \E[ \nu_0(X) \Delta Y | D=0] $$

. . .

The most common balancing weights are based on the propensity score, you can show:
\begin{align*}
  \nu_0(X) = \frac{p(X)(1-\pi)}{(1-p(X))\pi}
\end{align*}
where $p(X) = \P(D=1|X)$ and $\pi=\P(D=1)$.

* This is the approach suggested in @abadie-2005.  In practice, you need to estimate the propensity score.  The most common choices are probit or logit.

. . .

For estimation:

* Step 1: Estimate the propensity score (typically logit or probit)

* Step 2: Compute the weights, using the estimated propensity score

* Step 3: Compute $\widehat{ATT} = \displaystyle \frac{1}{n} \sum_{i=1}^n \frac{D_i}{\hat{\pi}} \Delta Y_i - \frac{1}{n} \sum_{i=1}^n \frac{\hat{p}(X_i) (1-D_i)}{\big(1-\hat{p}(X_i)\big) \hat{\pi}}$

## Augmented-Inverse Propensity Score Weighting (AIPW) {#doubly-robust}

You can show an additional identification result:
<!--$$ATT=\E\left[ \left( \frac{D}{p} - \frac{p(X)(1-D)}{(1-p(X))p} \right)(\Delta Y_{t} - \E[\Delta Y_{t} | X, D=0]) \right]$$-->

$$ATT = \E\left[  \Delta Y_{t} - \E[\Delta Y_{t} | X, D=0] \big| D=1\right] - \E\left[ \frac{p(X)(1-p)}{(1-p(X))p} \big(\Delta Y_{t} - \E[\Delta Y_{t} | X, D=0]\big) \Big| D = 0\right]$$


This requires estimating both $p(X)$ and $\E[\Delta Y|X,D=0]$.

. . .

Big advantage:  The sample analogue of this expression $ATT$ is [doubly robust]{.alert}.  This means that, it will deliver consistent estimates of $ATT$ if either the model for $p(X)$ or for $\E[\Delta Y|X,D=0]$ is correctly specified.

. . .

- In my experience, doubly robust estimators perform much better than either the regression or propensity score weighting estimators

. . .

- This also provides a connection to estimating $ATT$ under conditional parallel trends using machine learning for $p(X)$ and $\E[\Delta Y|X,D=0]$ (see: @chang-2020)


::: {.fragment}

[[more details](#understanding-double-robustness)]

:::

## Discussion

Regarding the previous issues with TWFE regressions, RA and AIPW satisfy:

[Issue 1:]{.alert} Multiple periods <span style="color:green">&#10004;</span>

[Issue 2:]{.alert} Time-invariant covariates <span style="color:green">&#10004;</span>

[Issue 3:]{.alert} Covariates affected by the treatment
<span style="color: blue; font-weight: bold;">?</span>

[Issue 4:]{.alert} Hidden linearity bias <span style="color:green">&#10004;</span>

[Issue 5:]{.alert} Weighted average of $ATT(X)$ <span style="color:green">&#10004;</span>

. . .

You can also show that they will, by construction, balance the means of $(X_{t=2},X_{t=1},Z)$ across groups.

. . .

In my view, these are much better properties that the TWFE regression when it comes to including covariates.


# Multiple Periods and Variation in Treatment Timing {visibility="uncounted"}

## Multiple Time Periods and Variation in Treatment Timing

<br>

::: {.callout-note}
### Conditional Parallel Trends with Multiple Periods

For all groups $g \in \bar{\mathcal{G}}$ (all groups except the never-treated group) and for all time periods $t=2, \ldots, T$,

$$\E[\Delta Y_{t}(0) | \mathbf{X}, Z, G=g] = \E[\Delta Y_{t}(0) | \mathbf{X}, Z, U=1]$$

where $\mathbf{X}_i := (X_{i1},X_{i2},\ldots,X_{iT})$.
:::

Under this assumption, using similar arguments to the ones above, one can show that

$$ATT(g,t) = \E\left[ \left( \frac{\indicator{G=g}}{\pi_g} - \frac{p_g(\mathbf{X},Z)U}{(1-p_g(\mathbf{X},Z))\pi_g}\right)\Big(Y_{t} - Y_{g-1} - m_{gt}^0(\mathbf{X},Z)\Big) \right]$$

where $p_g(\mathbf{X},Z) := \P(G=g|\mathbf{X},Z,\indicator{G=g}+U=1)$ and $m_{gt}^0(\mathbf{X},Z) := \E[Y_{t}-Y_{g-1}|\mathbf{X},Z,U=1]$.


## Practical Considerations

Because $\mathbf{X}_i$ contains $X_{i,t}$ for all time periods, terms like $m_{gt}^0(\mathbf{X},Z)$ can be quite high-dimensional (and hard to estimate) in many applications.

. . .

In many cases, it may be reasonable to replace with lower dimensional function $\mathbf{X}_i$:

. . .

* $\bar{X}_i$ &mdash; the average of $X_{it}$ across time periods
* $X_{it}, X_{ig-1}$ &mdash; the covariates in the current period and base period (this is possible in the `pte` package currently and may be added to `did` soon).
* $X_{ig-1}$ &mdash; the covariates in the base period (this is the default in `did`)
* $(X_{it}-X_{ig-1})$ &mdash; the change in covariates over time


. . .

Otherwise, however, everything is the same as before:

1. Recover $ATT(g,t)$

2. If desired: aggregate into $ATT^{es}(e)$ or $ATT^o$.

# Empirical Example {visibility="uncounted"}

::: {.notes}

* walk through a bunch of examples

* including covariates in the parallel trend assumtpion, compare to regression

* regression adjustment, ipw, doubly robust

* change base period

* change comparison group

* allow for anticipation

:::


## Back to Minimum Wage Example

Let's start by assuming that parallel trends holds conditional on a county's population and average income (sometimes we'll add region too)

* i.e., we would like to compare treated and untreated counties with similar populations and average incomes

. . .

I'll show results for the following cases:

1. Two period regression

2. All periods regression

    * with and without region as a covariate

3. Regression adjustment where the model is $Y_{it}(0) = \theta_t + \eta_i + X_{it}'\beta + e_{it}$

4. Callaway and Sant'Anna (2021) including $X_{g-1}$ and $Z$ as covariates

    * RA, IPW, AIPW

. . .

In addition to estimates, we'll also assess how well each of these works in terms of balancing covariates using the `twfeweights` package.



```{r}
#| echo: false
library(did)
library(BMisc)
library(twfeweights)
library(fixest)
library(modelsummary)
library(ggplot2)
load("data2.RData")
data2$region <- droplevels(data2$region)
```

## Two periods TWFE Regression

```{r}
# run TWFE regression
data2_subset <- subset(data2, year %in% c(2003,2004))
data2_subset <- subset(data2_subset, G %in% c(0, 2004))
twfe_x <- fixest::feols(lemp ~ post + lpop + lavg_pay | id + year,
                        data=data2_subset,
                        cluster="id")
modelsummary(twfe_x, gof_omit=".*")
```

## Diagnose covariate balance

```{r}
library(twfeweights)
tp_wts <- two_period_reg_weights(
  yname = "lemp",
  tname = "year",
  idname = "id",
  gname = "G",
  xformula = ~lpop + lavg_pay,
  extra_balance_vars_formula = ~region,
  data = data2_subset
)
```

## Diagnose covariate balance

```{r}
ggtwfeweights(tp_wts,
              absolute_value=FALSE,
              standardize=TRUE,
              plot_relative_to_target=FALSE) +
  xlim(c(-2,2))
```

## TWFE with more periods and covariates

```{r}
# run TWFE regression
twfe_x <- fixest::feols(lemp ~ post + lpop + lavg_pay | id + year,
                        data=data2,
                        cluster="id")
modelsummary(twfe_x, gof_omit=".*")
```

## Diagnose covariate balance

```{r}
#| cache: true
twfe_wts <- implicit_twfe_weights(
  yname = "lemp",
  tname = "year",
  idname = "id",
  gname = "G",
  xformula = ~lpop + lavg_pay,
  data = data2,
  base_period = "gmin1"
)
covariate_balance <- twfe_cov_bal(twfe_wts, ~ region + lpop + lavg_pay + -1)
```

## Diagnose covariate balance

```{r}
ggtwfeweights(covariate_balance,
              absolute_value = FALSE,
              standardize = TRUE,
              plot_relative_to_target = FALSE) +
  xlim(c(-1,1))
```

## Add region as a covariate

We'll allow for path of outcomes to depend on region of the country

```{r}
# run TWFE regression
twfe_x <- fixest::feols(lemp ~ post + lpop + lavg_pay | id + region^year,
                        data=data2,
                        cluster="id")
modelsummary(twfe_x, gof_omit=".*")
```

Relative to previous results, this is much smaller---this is (broadly) in line with the literature where controlling for `region` often matters a great deal (e.g., @dube-lester-reich-2010).

## Check covariate balance

```{r}
# similar code as before...check course materials
```

```{r}
#| echo: false
#| cache: true
twfe_wts2 <- implicit_twfe_weights(
  yname = "lemp",
  tname = "year",
  idname = "id",
  gname = "G",
  xformula = ~lpop + lavg_pay + as.factor(region)*as.factor(year),
  data = data2,
  base_period = "gmin1"
)
covariate_balance2 <- twfe_cov_bal(twfe_wts2, ~ region + lpop + lavg_pay + -1)

ggtwfeweights(covariate_balance2,
              absolute_value = FALSE,
              standardize = TRUE,
              plot_relative_to_target = FALSE) +
  xlim(c(-1,1))
```

## Regression adjustment with only $\Delta X$

```{r}
#| cache: true
# it's reg. adj. even though the function says aipw...
ra_wts <- implicit_aipw_weights(
  yname = "lemp",
  tname = "year",
  idname = "id",
  gname = "G",
  xformula = ~ 1,
  d_covs_formula = ~ lpop + lavg_pay,
  pscore_formula = ~1,
  data = data2
)
ra_wts$est
```

i.e., we estimate a somewhat larger effect of the minimum wage on teen employment

## Regression adjustment with only $\Delta X$

```{r}
ra_cov_bal <- aipw_cov_bal(ra_wts, ~ region + lpop + lavg_pay + -1)
ggtwfeweights(ra_cov_bal, absolute_value = FALSE,
              standardize = TRUE,
              plot_relative_to_target = FALSE) +
  xlim(c(-1,1))
```




## CS (2021) Regression Adjustment, $X_{g-1}, Z$


```{r eval=FALSE}
#| code-line-numbers: "|6|9"
# callaway and sant'anna including covariates
cs_x <- att_gt(yname="lemp",
               tname="year",
               idname="id",
               gname="G",
               xformla=~region + lpop + lavg_pay,
               control_group="nevertreated",
               base_period="universal",
               est_method="reg",
               data=data2)
cs_x_res <- aggte(cs_x, type="group")
summary(cs_x_res)
cs_x_dyn <- aggte(cs_x, type="dynamic")
ggdid(cs_x_dyn)
```


## CS (2021) Regression Adjustment, $X_{g-1},Z$

```{r echo=FALSE}
cs_x <- att_gt(yname="lemp",
               tname="year",
               idname="id",
               gname="G",
               xformla=~region + lpop + lavg_pay,
               control_group="nevertreated",
               base_period="universal",
               est_method="reg",
               data=data2)
cs_x_res <- aggte(cs_x, type="group")
summary(cs_x_res)
```

## CS (2021) Regression Adjustment, $X_{g-1},Z$

```{r echo=FALSE}
cs_x_dyn <- aggte(cs_x, type="dynamic")
ggdid(cs_x_dyn)
```

## Check covariate balance

```{r}
# similar code as before...check course materials
```


```{r}
#| echo: false
#| cache: true
ra_wts <- implicit_aipw_weights(
  yname = "lemp",
  tname = "year",
  idname = "id",
  gname = "G",
  xformula = ~ region + lpop + lavg_pay,
  d_covs_formula = ~ 1,
  pscore_formula = ~1,
  data = data2
)

ra_cov_bal <- aipw_cov_bal(ra_wts, ~ region + lpop + lavg_pay + -1)
ggtwfeweights(ra_cov_bal, absolute_value = FALSE,
              standardize = TRUE,
              plot_relative_to_target = FALSE) +
  xlim(c(-1,1))
```


## CS (2021) IPW, $X_{g-1}, Z$

```{r eval=FALSE}
#| code-line-numbers: "|9"
# callaway and sant'anna including covariates
cs_x <- att_gt(yname="lemp",
               tname="year",
               idname="id",
               gname="G",
               xformla=~region + lpop + lavg_pay,
               control_group="nevertreated",
               base_period="universal",
               est_method="ipw",
               data=data2)
cs_x_res <- aggte(cs_x, type="group")
summary(cs_x_res)
cs_x_dyn <- aggte(cs_x, type="dynamic")
ggdid(cs_x_dyn)
```



## CS (2021) IPW, $X_{g-1}, Z$

```{r echo=FALSE}
cs_x <- att_gt(yname="lemp",
               tname="year",
               idname="id",
               gname="G",
               xformla=~region + lpop + lavg_pay,
               control_group="nevertreated",
               base_period="universal",
               est_method="ipw",
               data=data2)
cs_x_res <- aggte(cs_x, type="group")
summary(cs_x_res)
```

## CS (2021) IPW, $X_{g-1}, Z$

```{r echo=FALSE}
cs_x_dyn <- aggte(cs_x, type="dynamic")
ggdid(cs_x_dyn)
```


## CS (2021) AIPW, $X_{g-1}, Z$

```{r eval=FALSE}
#| code-line-numbers: "|9"
# callaway and sant'anna including covariates
cs_x <- att_gt(yname="lemp",
               tname="year",
               idname="id",
               gname="G",
               xformla=~region + lpop + lavg_pay,
               control_group="nevertreated",
               base_period="universal",
               est_method="dr",
               data=data2)
cs_x_res <- aggte(cs_x, type="group")
summary(cs_x_res)
cs_x_dyn <- aggte(cs_x, type="dynamic")
ggdid(cs_x_dyn)
```



## CS (2021) AIPW, $X_{g-1}, Z$

```{r echo=FALSE}
cs_x <- att_gt(yname="lemp",
               tname="year",
               idname="id",
               gname="G",
               xformla=~region + lpop + lavg_pay,
               control_group="nevertreated",
               base_period="universal",
               est_method="dr",
               data=data2)
cs_x_res <- aggte(cs_x, type="group")
summary(cs_x_res)
```

## CS (2021) AIPW, $X_{g-1}, Z$

```{r echo=FALSE}
cs_x_dyn <- aggte(cs_x, type="dynamic")
ggdid(cs_x_dyn)
```

## Check covariate balance

```{r}
# similar code as before...check course materials
```


```{r}
#| echo: false
#| cache: true
aipw_wts <- implicit_aipw_weights(
  yname = "lemp",
  tname = "year",
  idname = "id",
  gname = "G",
  xformula = ~ region + lpop + lavg_pay,
  d_covs_formula = ~ 1,
  data = data2
)

aipw_cov_bal <- aipw_cov_bal(aipw_wts, ~ region + lpop + lavg_pay + -1)
ggtwfeweights(aipw_cov_bal, absolute_value = FALSE,
              standardize = TRUE,
              plot_relative_to_target = FALSE) +
  xlim(c(-1,1))
```

## Discussion

If you want to include covariates in the parallel trends assumption, it is better to use approaches that directly include the covariates relative to estimation strategies that transform the covariates

# Covariates Affected by the Treatment {#bad-controls-details visibility="uncounted"}

## Covariates Affected by the Treatment {#covariates-affected-by-the-treatment}

So far, our discussion has been for the case where the time-varying covariates [evolve exogenously]{.alert}.

* Many (probably most) covariates fit into this category: in the minimum wage example, a county's population probably fits here.

. . .

In some applications, we may want to control for covariates that themselves could be affected by the treatment

* Classical examples in labor economics: A person's industry, occupation, or union status

* These are often referred to as ["bad controls"]{.alert}

. . .

You can see a [tension]{.alert} here:

* We would like to compare units who, absent being treated, would have had the same (say) union status

* But union status could be affected by the treatment

## Covariates Affected by the Treatment

The most common practice is to just completely drop these covariates from the analysis

* Not clear if this is the right idea though...

We will consider some alternatives

* Condition on pre-treatment value of bad control
* Treat bad control as an outcome (i.e., use some identification strategy), then feed this into the main analysis as a covariate


## Additional Notation

To wrap our heads around this, let's go back to the case with two time periods.

Define treated and untreated [potential covariates]{.alert}: $X_{it}(1)$ and $X_{it}(0)$.  Notice that in the "textbook" two period setting, we observe
$$X_{i,t=2} = D_i X_{i,t=2}(1) + (1-D_i) X_{i,t=2}(0) \qquad \textrm{and} \qquad X_{i,t=1} = X_{i,t=1}(0)$$

Then, we will consider parallel trends in terms of untreated potential outcomes and untreated potential covariates:

<br>

::: {.callout-note}
### Conditional Parallel Trends using Untreated Potential Covariates

$$\E[\Delta Y(0) | X_{t=2}(0), X_{t=1}(0), Z, D=1] = \E[\Delta Y(0) | X_{t=2}(0), X_{t=1}(0), Z, D=0]$$
:::

<br>

## Identification Issues

Following the same line of argument as before, it follows that

$$ATT = \E[\Delta Y | D=1] - \E\Big[ \E[\Delta Y(0) | X_{t=2}(0), X_{t=1}(0), Z, D=0] \Big| D=1\Big]$$

The second term is the tricky one.  Notice that:

* The inside conditional expectation is identified &mdash; we see untreated potential outcomes and covariates for the untreated group

* However, we cannot average over $X_{t=2}(0)$ for the treated group, because we don't observe $X_{t=2}(0)$ for the treated group

. . .

There are several options for what we can do &nbsp; $\rightarrow$

## Option 1: Ignore

One idea is to just ignore that the covariates may have been affected by the treatment:

<br>

::: {.callout-note}

### Alternative Conditional Parallel Trends 1

$$\E[\Delta Y(0) | { \color{red} X_{\color{red}{i,t=2} } }, X_{t=1}(0), Z, D=1] = \E[\Delta Y(0) |  { \color{red} X_{\color{red}{i,t=2}} }, X_{t=1}(0), Z, D=0]$$
:::

<br>

The limitations of this approach are well known (even discussed in MHE), and this is not typically the approach taken in empirical work

. . .

[Job Displacement Example:]{.alert-blue} You would compare paths of outcomes for workers who left union [because they were displaced]{.alert} to paths of outcomes for non-displaced workers who also left union (e.g., [because of better non-unionized job opportunity]{.alert})

## Option 2: Drop

It is more common in empirical work to drop $X_{i,t}(0)$ entirely from the parallel trends assumption

<br>

::: {.callout-note}

### Alternative Conditional Parallel Trends 2

$$\E[\Delta Y(0) | Z, D=1] = \E[\Delta Y(0) |  Z, D=0]$$
:::

<br>

In my view, this is not attractive either though.  If we believe this assumption, then we have basically solved the bad control problem by assuming that it does not exist.

. . .

[Job Displacement Example:]{.alert-blue} We have now just assumed that path of earnings (absent job displacement) doesn't depend on union status

## Option 3: Tweak

Perhaps a better alternative identifying assumption is the following one

<br>

::: {.callout-note}

### Alternative Conditional Parallel Trends 3

$$\E[\Delta Y(0) | X_{t=1}(0), Z, D=1] = \E[\Delta Y(0) |  X_{t=1}(0), Z, D=0]$$
:::

<br>

. . .

[Intuition:]{.alert} Conditional parallel trends holds after conditioning on pre-treatment time-varying covariates that could have been affected by treatment

. . .

[Job Displacement Example:]{.alert-blue} Path of earnings (absent job displacement) depends on pre-treatment union status, but not untreated potential union status in the second period

. . .

[What to do:]{.alert} Since $X_{i,t=1}(0)$ is observed for all units, we can immediately operationalize this assumption use our arguments from earlier (i.e., the ones without bad controls)

* This is difficult to operationalize with a TWFE regression

* In practice, you can just include the bad control among other covariates in `did`

::: {.notes}

* It is not the same identifying assumption as the one that we started with, but we are at least allowing for the bad control to show up in the identifying assumption

:::

## Option 4: Extra Assumptions

Another option is to keep the original identifying assumption, but add additional assumptions where we (in some sense) treat $X_t$ as an outcome and as a covariate.

Recall:

$$ATT = \E[\Delta Y | D=1] - \E\Big[ \E[\Delta Y(0) | X_{t=2}(0), X_{t=1}(0), Z, D=0] \Big| D=1\Big]$$

. . .

If we could figure out distribution of $X_{t=2}(0)$ for the treated group, we could recover $ATT$

## Option 4: Dealing with $X_{t=2}(0)$

::: {.callout-note}

### Covariate Unconfoundedness Assumption

$$X_{t=2}(0) \independent D | X_{t=1}(0), Z$$

:::

[Intuition:]{.alert} For the treated group, the time-varying covariate would have evolved in the same way over time as it actually did for the untreated group, conditional on $X_{t=1}$ and $Z$.

* Notice that this assumption only concerns untreated potential covariates $\implies$ it allows for $X_{t=2}$ to be affected by the treatment

* Making an assumption like this indicates that $X_{t=2}(0)$ is playing a dual role: (i) start by treating it as if it's an outcome, (ii) have it continue to play a role as a covariate

Under this assumption, can show that we can recover the $ATT$:

$$ATT = \E[\Delta Y | D=1] - \E\left[ \E[\Delta Y | X_{t=1}, Z, D=0] \Big| D=1 \right]$$

This is the same expression as in Option 3

## Option 4: Additional Discussion

In some cases, it may make sense to condition on other additional variables (e.g., the lagged outcome $Y_{t=1}$) in the covariate unconfoundedness assumption.  In this case, it is still possible to identify $ATT$, but it is more complicated

It could also be possible to use alternative identifying assumptions besides covariate unconfoundedness &mdash; at a high-level, we somehow need to recover the distribution of $X_{t=2}(0)$

* e.g., @brown-butts-westerlund-2023

See @caetano-callaway-payne-santanna-2022 for more details about bad controls.

[[Back](#covariates-affected-by-the-treatment)]



# Appendix {visibility="uncounted"}

## Regression weights - more details {#reg-weights-more-details visibility="uncounted"}

By construction, the regression weights balance the mean of $X$.  That is,
$$ \E\Big[ w_1(X) X \Big| D=1\Big] = \E\Big[ w_0(X) X \Big| D=1 \Big] $$

. . .

But they do not necessarily balance other functions of the covariates such as quadratic terms, interactions, etc.  You can check for balance by computing terms like
$$ \E\Big[ w_1(X) X^2 \Big| D=1\Big] \overset{?}{=} \E\Big[ w_0(X) X^2 \Big| D=1 \Big] $$

. . .

(perhaps) that the regression weights balance the means of covariates indicates that typically the misspecification bias should be small

[[Back](#interpreting-regressions)]


## Understanding Double Robustness {#understanding-double-robustness visibility="uncounted"}

To understand double robustness, we can rewrite the expression for $ATT$ as
\begin{align*}
  ATT = \E\left[ \frac{D}{\pi} \Big(\Delta Y - m_0(X)\Big) \right] - \E\left[ \frac{p(X)(1-D)}{(1-p(X))\pi} \Big(\Delta Y - m_0(X)\Big)\right]
\end{align*}

The first term is exactly the same as what comes from regression adjustment

* If we correctly specify a model for $m_0(X)$, it will be equal to $ATT$.

* If $m_0(X)$ not correctly specified, then, by itself, this term will be biased for $ATT$

The second term can be thought of as a de-biasing term

* If $m_0(X)$ is correctly specified, it is equal to 0

* If $p(X)$ is correctly specified, it reduces to $\E[\Delta Y_{t}(0) | D=1] - \E[m_0(X)|D=1]$ which both delivers counterfactual untreated potential outcomes and removes the (possibly misspecified) second term from the first equation

[[Back](#doubly-robust)]


# References {visibility="uncounted"}

::: {#refs}
:::